Linking hypothesis and number of response options modulate inferred scalar implicature rate

### 2018-05-10 manuscript.tex
final edits from jd

Let's examine an example. Assume three response options (wrong, neither,
right). Assume further that each of the three responses was selected by
a third of participants, i.e., the distributions of responses is 1/3,
1/3, and 1/3. Under the Strong link, we infer that this task yielded an
implicature rate of 2/3. Under the Weak link, we infer that this task
yielded an implicature rate of 1/3. This is quite a drastic difference
if we are, for instance, interested in whether scalar implicatures are
inference defaults and we would like to interpret an implicature rate of
above an arbitrary threshold (e.g., 50\%) as evidence for such a claim.
Under the Strong link, we would conclude that scalar implicatures are
not defaults. Under the Weak link, we would conclude that they are. In
the experiment reported in the following section, we presented
participants with exactly this setup. We manipulated the number of
response options between participants and analyzed the results under
different linking hypothesiss.

@@

Participants played an online card game in which they were asked to
judge descriptions of the contents of cards. Different groups of
participants were presented with different numbers of response options.
On critical trials, participants were presented with descriptions for
the cards that typically result in exhaustivity implicatures
(\enquote{There is a cat on the card} when there was a cat and a dog) or
scalar implicatures (\enquote{There is a cat or a dog on the card} when
there was a cat and a dog). We categorized their responses on such
trials according to the weak and the Strong link introduced above, and
tested whether the number of response options and the linking
hypothesiss led to different conclusions about the rate of computed
implicatures in the experimental task.

@@

The study was administered online through Amazon Mechanical
Turk.\footnote{The experiment can be viewed
  \href{https://cdn.rawgit.com/thegricean/si-paradigms/94a590f0/experiments/main/1_methods/online_experiment/connective_game.html}{here}.}
Participants were first introduced to the set of cards we used in the
study (Figure \ref{fig:stimuli}). Each card depicted one or two animals,
where an animal could be either a cat, a dog, or an elephant. Then
participants were introduced to a blindfolded fictional character called
Bob. Bob was blindfolded to avoid violations of ignorance expectations
associated with the use of disjunction (Chierchia et al., 2001;
Sauerland, 2004) Participants were told that Bob would guess the
contents of the cards and their task was to indicate whether Bob's guess
was wrong or right. On each trial, participants saw a card and a
sentence representing Bob's guess. For example, they saw a card with a
cat and read the sentence \enquote{There is a cat on the card.} They
then provided an assessment of Bob's guess. The study ended after 24
trials.

@@

In this paper we asked whether linking hypothesiss and number of
response options available to participants in truth value judgment tasks
affects inferred implicature rates. The results presented here suggest
they do. A linking assupmtion that considered the highest point on the
scale literal and any lower point pragmatic (Weak link) resulted in
higher implicature rates in tasks with 4 or 5 response options compared
to the standard two options. A linking hypothesis that considered the
lowest point on the scale pragmatic and any higher point literal (Strong
link) reported lower implicature rates in tasks with 4 or 5 options
compared to the standard two options. The results suggest that the
choice of linking hypothesis is a crucial analytical step that can
significantly impact the conclusions drawn from truth value judgment
tasks. In particular, there is danger for pragmatic ability to be both
under- and overestimated.

While the binary truth value judgement task avoids the analytic decision
between Strong and Weak linking hypothesiss, the results reported here
suggest that binary tasks can also underestimate participants' pragmatic
competence. In binary tasks, participants are often given the lowest and
highest points on a scale (\enquote{wrong} vs. \enquote{right}) and are
asked to report pragmatic infelicities using the lowest point (e.g.
\enquote{wrong}). The study reported here showed that on trials with
true but pragmatically infelicitous descriptions, participants often
avoided the lowest point on the scale if they were given more
intermediate options. Even though the option \enquote{wrong} was
available to participants in all tasks, participants in tasks with
intermediate options chose it less often. In computing implicature rate,
this pattern manifested itself as a decrease in implicature rate under
the Strong link when more response options were provided, and an
increase in implicature rate under the Weak link when more response
options were provided. These observations are in line with Katsos and
Bishop (2011)\enquote{s argument that pragmatic violations are not as
severe as semantic violations and participants do not penalize them as
much. Providing participants with only the extreme ends of the scale
(e.g.~wrong/right, false/true) when pragmatic violations are considered
to be of an intermediate nature risks misrepresentation of participants}
pragmatic competence. It further suggests that in studies that use
binary tasks to investigate response-contingent processing, proportions
of \enquote{literal} responses may be a composite of both literal and
pragmatic underlying interpretations that just happen to get mapped
differently onto different response options by participants.

@@

\subsection{Revisiting linking
hypothesiss}\label{revisiting-linking-hypothesiss}

@@

Concluding, we have shown in this paper that inferred
\enquote{implicature rate} -- ubiquitous in theoretical and experimental
pragmatics -- as estimated in truth value judgment tasks, depends on
both the number of responses participants are provided with as well as
on the linking hypothesis from proportion of behavioral responses to
\enquote{implicature rate}. We further sketched an alternate linking
hypothesis that treats behavioral responses as the result of
probabilistic reasoning about speakers' likely productions. While a
thorough model comparison is still outstanding, this kind of linking
hypothesis opens a door towards more systematic and rigorous formulation
and testing of linking hypotheses between theoretical notions of
interest in pragmatics and behavioral responses in experimental
paradigms.

>>>>

Let's examine an example. Assume three response options (wrong, neither,
right). Assume further that each of the three responses was selected by
a third of participants, i.e., the distributions of responses is 1/3,
1/3, and 1/3. Under the Strong link, we infer that this task yielded an
implicature rate of 2/3. Under the Weak link, we infer that this task
yielded an implicature rate of 1/3. This is quite a drastic difference
if we are, for instance, interested in whether scalar implicatures are
inference defaults and we would like to interpret an implicature rate of
above an arbitrary threshold (e.g., 50\%) as evidence for such a claim.
Under the Strong link, we would conclude that scalar implicatures are
not defaults. Under the Weak link, we would conclude that they are. In
the experiment reported in the following section, we presented
participants with exactly this setup. We manipulated the number of
response options between participants and analyzed the results under
different linking hypothesis.

@@

Participants played an online card game in which they were asked to
judge descriptions of the contents of cards. Different groups of
participants were presented with different numbers of response options.
On critical trials, participants were presented with descriptions for
the cards that typically result in exhaustivity implicatures
(\enquote{There is a cat on the card} when there was a cat and a dog) or
scalar implicatures (\enquote{There is a cat or a dog on the card} when
there was a cat and a dog). We categorized their responses on such
trials according to the Weak and the Strong link introduced above, and
tested whether the number of response options and the linking
hypothesis led to different conclusions about the rate of computed
implicatures in the experimental task.

@@

The study was administered online through Amazon Mechanical
Turk.\footnote{The experiment can be viewed
  \href{https://cdn.rawgit.com/thegricean/si-paradigms/94a590f0/experiments/main/1_methods/online_experiment/connective_game.html}{here}.}
Participants were first introduced to the set of cards we used in the
study (Figure \ref{fig:stimuli}). Each card depicted one or two animals,
where an animal could be either a cat, a dog, or an elephant. Then
participants were introduced to a blindfolded fictional character called
Bob. Bob was blindfolded to avoid violations of ignorance expectations
associated with the use of disjunction (Chierchia et al., 2001;
Sauerland, 2004). Participants were told that Bob would guess the
contents of the cards and their task was to indicate whether Bob's guess
was wrong or right. On each trial, participants saw a card and a
sentence representing Bob's guess. For example, they saw a card with a
cat and read the sentence \enquote{There is a cat on the card.} They
then provided an assessment of Bob's guess. The study ended after 24
trials.

@@

In this paper we asked whether linking hypothesis and number of
response options available to participants in truth value judgment tasks
affects inferred implicature rates. The results presented here suggest
they do. A linking assupmtion that considered the highest point on the
scale literal and any lower point pragmatic (Weak link) resulted in
higher implicature rates in tasks with 4 or 5 response options compared
to the standard two options. A linking hypothesis that considered the
lowest point on the scale pragmatic and any higher point literal (Strong
link) reported lower implicature rates in tasks with 4 or 5 options
compared to the standard two options. The results suggest that the
choice of linking hypothesis is a crucial analytical step that can
significantly impact the conclusions drawn from truth value judgment
tasks. In particular, there is danger for pragmatic ability to be both
under- and overestimated.

While the binary truth value judgement task avoids the analytic decision
between Strong and Weak linking hypothesis, the results reported here
suggest that binary tasks can also underestimate participants' pragmatic
competence. In binary tasks, participants are often given the lowest and
highest points on a scale (\enquote{wrong} vs. \enquote{right}) and are
asked to report pragmatic infelicities using the lowest point (e.g.
\enquote{wrong}). The study reported here showed that on trials with
true but pragmatically infelicitous descriptions, participants often
avoided the lowest point on the scale if they were given more
intermediate options. Even though the option \enquote{wrong} was
available to participants in all tasks, participants in tasks with
intermediate options chose it less often. In computing implicature rate,
this pattern manifested itself as a decrease in implicature rate under
the Strong link when more response options were provided, and an
increase in implicature rate under the Weak link when more response
options were provided. These observations are in line with Katsos and
Bishop (2011)\enquote{s argument that pragmatic violations are not as
severe as semantic violations and participants do not penalize them as
much. Providing participants with only the extreme ends of the scale
(e.g.~wrong/right, false/true) when pragmatic violations are considered
to be of an intermediate nature risks misrepresentation of participants}
pragmatic competence. It further suggests that in studies that use
binary tasks to investigate response-contingent processing, proportions
of \enquote{literal} responses may be a composite of both literal and
pragmatic underlying interpretations that just happen to get mapped
differently onto different response options by participants.

@@

\subsection{Revisiting the linking
hypothesis}\label{revisiting-linking-hypothesis}

@@

Concluding, we have shown in this paper that inferred
\enquote{implicature rate} -- a ubiquitous notion in theoretical and experimental
pragmatics -- as estimated in truth value judgment tasks, depends on
both the number of responses participants are provided with as well as
on the linking hypothesis from proportion of behavioral responses to
\enquote{implicature rate}. We further sketched an alternate linking
hypothesis that treats behavioral responses as the result of
probabilistic reasoning about speakers' likely productions. While a
thorough model comparison is still outstanding, this kind of linking
hypothesis opens a door towards more systematic and rigorous formulation
and testing of linking hypotheses between theoretical notions of
interest in pragmatics and behavioral responses in experimental
paradigms.
